---
title: "Titanic Data Wrangling Exercise"
author: "Laura Rose"
date: "May 23, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load libraries

```{r}
library(tidyr)
library(dplyr)
```
## Step 0: Load the Data in RStudio

```{r}
titanic_original <- read.csv("C:/Users/laura.rose/Documents/titanic_original.csv")
```

## Step 1: Port of Embarkation

replace missing values in embarked variable 


```{r}
titanic_original$embarked <- with(titanic_original, replace(embarked, embarked=="", "S"))
```

## Step 2: Age
```{r}
mean_age <- mean(titanic_original$age, na.rm = TRUE)
titanic_original$age <- replace(titanic_original$age, is.na(titanic_original$age), mean_age)
```
Another replacement for the missing values could be the median. This might be more representative of the passengers since the mean is affected by outliers.

## Step 3: Lifeboat
```{r}
titanic_original <- mutate(titanic_original, boat = ifelse(trimws(titanic_original$boat)=="", "None", boat))
```

## Step 4: 
It would not be useful to replace a missing value for a cabin number because the actual combination of numbers and letters for a given cabin is not important, but rather whether it is known which cabin a passenger had. A missing value could imply that the passenger didn't survive, so nobody after the sinking of the ship knew their cabin number.

```{r}
titanic_original <- mutate(titanic_original, has_cabin_number = as.numeric(trimws(titanic_original$cabin)!=""))
```

## Step 5: Write the cleaned-up data to a CSV

```{r}
write.csv(titanic_original, "C:/Users/laura.rose/Documents/titanic_clean.csv")
```